---

title: "P8131_hw7"
author: Shan Jiang 
output:
     pdf_document:
         latex_engine: xelatex
fontsize: 10.5pt
geometry: margin=0.6in 

---


#### Import data 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lme4)
library(nlme)
health  = readxl::read_xlsx("./HW8-HEALTH.xlsx")
head(health)
## factorize the variable HEALTH, TXT 
health$HEALTH = as.factor(health$HEALTH)
levels(health$HEALTH)

health$TXT = as.factor(health$TXT)
levels(health$TXT)

##  n=80:  4 points in time: randomization, 3 months, 6 months, and 12 months post-randomization.
```

## (a) Baseline comparison 

### Boxplots to show the relationship

#### 1. Descriptive statistics

```{r}
## subset data which only have value at the baseline level 
health.sub <- subset(health, TIME == "1")
nrow(health.sub)

## There are in total 80 baseline individual rows in our dataset
## Barplot 
health.sub %>% 
  group_by(TXT, HEALTH) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x = TXT, y = n, fill = HEALTH)) +  
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = c("Intervention","Control"))

```


From this stacked barplot, we can see a slight difference in the baseline status between the intervention group and control group, with control group has a higher poor health status subjects in their pool while the intervention group subjects are more balanced in their health status.  


#### 2. Quantatitive comparison

```{r}
# Fit a simple GLM model 
hea_glm <- glm(HEALTH ~ TXT, data = health.sub , family = "binomial")
summary(hea_glm)

```

The GLM model shows that the p-value is 0.486, which is greater than 0.05, so there is no significant difference between the control group and intervention group in their subjects' health status. 



## (b) GEE model

```{r message=TRUE, warning=TRUE, paged.print=TRUE}
library(gee) # Note: data need to be sorted!!! make sure measures from the same subject are together!
dim(health) # there are 279 datapoints, and it is an unbalanced design as there are some data points missing
```

### reconstruct the variable: baseline status and month 

```{r}
# make time 1 as another covariate: baseline status
health = health %>%
  group_by(ID) %>% 
  mutate(HEALTH = as.numeric(HEALTH == "Good") ) %>% 
   mutate(baseline = ifelse(TIME == "1", HEALTH[TIME =="1"], HEALTH[TIME =="1"])) %>% 
   mutate(treatment = TXT) %>% 
   select(-TXT)

health$AGEGROUP = as.factor(health$AGEGROUP)
levels(health$AGEGROUP) 

health$baseline = recode(health$baseline, "1" = "0", "0" = "1")


### month post randomization: months
health = health %>% 
  mutate(month = if_else(TIME == "4", (TIME - 1) * 4, (TIME - 1) * 3))

```


### Fit model GEE.

```{r}
# Unstructured 
health.follow <- subset(health, month > "0")

hea_gee1 <- gee(HEALTH ~ month + treatment + baseline + AGEGROUP, data = health.follow, 
                 family = "binomial", id = ID,
                 corstr = "unstructured", 
                 scale.fix = TRUE, scale.value = 1) # scale parameter is phi (over dispersion)

summary(hea_gee1)
```

#### Exponentiate values back. 

```{r}
## values reflect effects on the log-odds scale
exp(coef(hea_gee1)["month"])
exp(coef(hea_gee1)["treatmentIntervention"])  
exp(coef(hea_gee1)["baseline"]) 
exp(coef(hea_gee1)["AGEGROUP25-34"]) 
exp(coef(hea_gee1)["AGEGROUP35+"]) 
```


* GEE model:

* Interpret your results:

* Intercept: the log odds of good health rating is 0.12457924 on average for patients with Good baseline health rating and age within 15-24 at control group at baseline.

* $\beta{Month}$: For patients within the same age group and same treatment group and same baseline rating, one unit increase of month is associated with 0.03243 increase in log odds ratio in being at good health status when keeping all variables constant.

* $\beta_{Age25-34}$: The expected value of log odds ratio is 1.3525 being at good health status in the AGE GROUP of 25-34 vs these who do not fall into the age group, adjusted for all other covariates for patients within the same treatment group and same baseline rating at same post randomization month.

* $\beta_{AGEGROUP35+}$: : For patients within the same treatment group and same baseline rating at same post randomization month, the log odds ratio of good health rating for age above 35 vs. age within 15-24 is 1.42052166 on average, adjusted for all other covariates. 

* $\beta_{Treatment}$: The expected value of log odds ratio for being at good health status is  2.10226 for these who have intervention vs these who are of control, adjusted for all other covariates. 

* $\beta_{baseline1}$: The expected value of log odds ratio is -1.8141 of being at good health status for subject who are of good health status at the baseline compared with these who are of poor health status at the baseline, adjusted for all other covariates. 



(c) Fit a GLMM.

```{r}
# fit GLMM 
# random intercept model

h.GLMM1 <- glmer(HEALTH  ~ baseline + AGEGROUP +  treatment + month + (1 | ID), 
                  family = 'binomial', data =  health.follow)

summary(h.GLMM1) 

# correlation of fixed effects is related to Fisher information of estimates
random.effects(h.GLMM1)

fixed.effects(h.GLMM1)

```

$$g(E(Y_{ij}|b_i)) = X_{ij}\beta + Z_{ij}b_i+ \epsilon_{ij}$$

Interpretation: 

* Only the time variable $month$ can be interpreted in this experiement design. 

* $month$: For patients within the same age group and same treatment group and same baseline rating, the log odds ratio of good health rating for one month increase is 0.03718.

* bashealthPoor, agegroup25-34, agegroup35+ cannot be interpreted, because for the same subject, these characteristics are not interchangeable, and there is no point in comparing it to different levels.


*The difference is that the GEE model* focuses on modelling Population-average while the GLMM model focuses on the specific individual and some variables cannot be changed for an individual.



