---
title: "P8131_hw10"
author: Shan Jiang 
output:
     pdf_document:
         latex_engine: xelatex
fontsize: 11.5pt
geometry: margin=1.1in 
---

```{r include=FALSE}
library(flexsurv)
library(survival)
library(MASS)
library(survminer)
```


## Problem 1

```{r}
# load data 
#############################################
time  = c(4, 2, 12, 6, 15, 8, 21, 10, 23, 19)
pair = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5)
cens = c(0, 0, 1, 1, 0, 1, 1, 0, 0, 0)
treat = rep(c(1, 2), 5 )

df = data.frame(pair, time, cens, treat)
head(df)
```


$H_0: h_1(t) = h_2(t)$ for all t
$H_1: h_1(t) \neq h_2(t)$ 

```{r}
## Log Rank test
survdiff(Surv(time,cens)~treat, data = df) # log rank test

ggsurvplot( survfit(Surv(time, cens) ~ treat, data = df), conf.int = TRUE)
```

Interpretation: 

* The log rank test: at the significance of 0.05, we cannot reject the null hypothesis as p > 0.3.So, we cannot say there is significant difference for hazard function between these two groups.


## Problem 2

```{r}
library(KMsurv)
data(kidtran)
```

 * gender: 1 = Male, 2 = Female;
 * race: 1 = white, 2 = black; 
 * delta: Death indicator (0 = alive, 1 = dead);

### Two gender group data: visual comparison

#### Group 1: Male(gender = 1)

```{r}
## gender: Male
ggsurvplot(survfit(Surv(time, delta) ~  race, data = subset(kidtran, gender == "1")), conf.int=TRUE)

```

More formal way to compare intra-group differences, is by using the log-rank test to test hypotheses.

```{r}
# Male
survdiff(Surv(time,delta) ~ race, data = subset(kidtran, gender == "1"))
```

In the log-rank test, we fail to reject the null hypothesis and conclude that there is no significant difference in two races groups for t defined. 

```{r}
fit = coxph(Surv(time, delta) ~  gender + race, 
            data = kidtran)
summary(fit)

## In male group 
fit_male = coxph(Surv(time, delta) ~ race, 
            data = subset(kidtran, gender == "1"))
summary(fit_male)
```

#### Group 2: Female(gender = 2)

```{r}
## gender: Female 
ggsurvplot(survfit(Surv(time, delta) ~  race, data = subset(kidtran, gender == "2")), conf.int=TRUE)

# Female
survdiff(Surv(time,delta) ~ race, data = subset(kidtran, gender == "2"))
```

For female subjects, there is a significant difference in hazards ratio (p < 0.03) for Black and white people.

```{r}
## Female(gender = 2)
fit_male = coxph(Surv(time, delta) ~ age + race, 
            data = subset(kidtran, gender == "2"))
summary(fit)
```


## Problem 3

```{r}
data(larynx)

larynx = larynx %>%  
            mutate( Z1 = ifelse(stage == 2, 1, 0)) %>% 
            mutate( Z2 = ifelse(stage == 3, 1, 0)) %>% 
                mutate( Z3 = ifelse(stage == 4, 1, 0)) %>% 
                mutate( Z4 = age) %>% 
                mutate( Z1_Z4 =  Z1 * Z4) # ADD interaction term Z1 * Z

head(larynx)
```


#### Model building 

$h(t) = h_0(t) exp(\beta_1Z_1 + \beta_2Z_2 + \beta_3Z_3 + \beta_4Z_4 + \beta_5Z_1 \times Z_4$

```{r}

fit_3 = coxph(Surv(time, delta) ~  age + Z1 + Z2 + Z3 + Z4 + Z1_Z4, 
            data = larynx, ties = 'breslow')

summary(fit_3)

```

Interpretation of results:

* $\beta_1 + \beta_5$: The log hazard ratio for subjects in stage II versus stage I is (-7.382 + 0.112 $\times$age) given of the same age.

* $\beta_2$: The log hazards ratio for subjects in stage III versus stage I is 0.621 given the same age.

* $\beta_3$: The log hazards ratio for subjects in stage IV versus stage I is 1.753 given they have same age.

* $\beta_4$: Compared to non-Stage II, the log hazards ratio for subjects with one unit changes in age is 0.006 given they are in same stage.

* $\beta_4 + \beta_5$: In the stage II for subjects, the log hazards ratio for subjects with one unit
changes in age is 0.118.

Relative risk:

* For the hazards of dying for a stage II patient of age 50 is:
$$h_2(t) = h_0(t) \times exp(−7.382 × 1 + 0.111 × 1 × 5)$$

* For the hazard of dying for a stage I patient of age 50 is
$$h_1(t) = h_0(t)exp(0)$$
* *So* the hazard ratio: $HR(t) = \frac{h_2(t)}{h_1(t)} = 0.16$
