---
title: "P8131_hw5"
author: Shan Jiang 
output:
     pdf_document:
         latex_engine: xelatex
fontsize: 10.5pt
geometry: margin=0.8in
---

### Problem  1 nesting crabs 

```{r message=FALSE, warning=FALSE, include=FALSE}
## 1.Active packages
library(pscl)
library(ggplot2)
library(tidyverse)
library(stargazer)
library(formattable)
```

#### a). fit a poisson model M1 with logit link W.

```{r fig.height=4.5, fig.width=3.5, message=FALSE, warning=FALSE}
# import in the crab data
crab = read.table('./HW5-crab.txt', header = T)
# test if there is obs = 0 for W and Wt, no 0 exists for model 
nrow(crab)
nrow(subset(crab, crab$W > 0)) 

crab.M1 <- glm(crab$Sa~W, family = poisson(link = log), data = crab)
summary(crab.M1) 
# n = 173, beta0 =-3.30476, beta1 = 0.16405 
anova(crab.M1)
```

The response outcome for each female crab is her number of satellites (Sa), with predictor as carapace width (W), estimated model is as following:
$$log(\hat {Sa}_i) = 0.16405*W_i -3.30476$$
The Std.Error of estimated $\beta_1$ = 0.164 is 0.01997 which is small, and the slope is statistically significant given its z-value of 8.216 and its low p-value.

Interpretation: 
Since estimate of $\beta_1$> 0, for one unit of increase in the female crab width, the expected number of male satellites will increase, and will be multiplied by $e^{\beta_1}$ = 1.18 times.

#### Goodness of fit

(1) Graphs test.
```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE}
## we may suspect some outliers from the simple Y-X scatter plot;
ggplot(data = crab, aes(x = W, y = Sa )) +
  geom_point(shape = 1)  +
  geom_smooth(method = glm)
## fit the log model:
ggplot(data = crab, aes(x = log(W + 1), y = Sa )) +
  geom_point(shape = 1)  +
  geom_smooth(method = glm)

```

From two graphs, it looks like there is no significant difference for the log model compared with the simple linear regression model, we may consider more about the outliers aligned in the bottom part and the top grid of the panel.

(2) Model diagnostic: Goodness of Fit

```{r}
## Methods 1: Genearalized Pearson Chi-square
res.p1 = residuals(crab.M1, type = 'pearson',data = crab)  
G1 = sum(res.p1^2) # calc dispersion param based on full model
## [1] 544.157

df = 173 - 2 
1 - pchisq(G1, df = df) ## 
## [2] df = 171, p-value = 0 


## Alternative Methods: Deviance Analysis
D = deviance(crab.M1)
## [1] 567.8786
1 - pchisq(D, df = df) 
## [2] p-value = 0, means 
```

The Pearson's chi-squared and deviance both are around 540 and 560, and maintained a p-value of 0 < 0.05, meaning this model doesn't fit well, we need an adjustment.

#### b). fit a poisson model M2

```{r}
## There is offset in this dataset
crab.M2 <- glm(Sa ~ Wt + W, family = poisson(link = log), data = crab)
summary(crab.M2) 
```
After adding a new predictor Wt, we have a larger model:
The response outcome for each female crab is her number of satellites (Sa), with predictors as carapace width (W), and weight(Wt) in estimated model is as following:
$$log(\hat {Sa}_i) = 0.44744W_t + 0.04590 * W -1.29168$$

Interpretation: 
* Since estimate of $\beta_1$> 0, for one unit of increase in the female crab weight, the expected number of male satellites will increase, and will be multiplied by $e^{\beta_1}$ = 1.5643 times.

* Although estimate of $\beta_2$ is also positive, the coefficient didn't past the Z-test which indicates that there is no significant influence of female crab width on the expected number of male satellites.

#### Model comparison:

Since the model M1 is nested in model M2, we can have a deviance analysis for comparing these two models.
```{r}
D.stat = crab.M1$deviance - crab.M2$deviance
df.test =  171 - 170 
pval.test = 1 - pchisq(D.stat, df = df.test) # chisq test
pval.test 
```

df for M1: 173 - 2 = 171
df for M2: 173 - 3 = 170

Compared with model M1, the coefficient for female crab width became smaller and not significant at all(p-value = 0.32), which alters the original conclusion, while the weight of male crab significantly affects the number of satellites in the M2 model. Because the p-value is 0.00469 < 0.05, so we need to reject the null hypothesis, and say that the M2 model is a better fit than M1. 


### c). Check the over-dispersion. 

(1) estimate the dispersion parameter 

```{r}
# the traditional way of calc constant dispersion parameter
res.p2 = residuals(crab.M2, type = 'pearson', data = crab)  
G1 = sum(res.p2^2)
# Generalized-chi-square: 536.5963

## Estimate the dispersion parameter
phi = G1/(173 - 3) 
phi # 3.15644

crab.M2$deviance/crab.M2$df.residual 
# 3.293442
```

(2) Half-normal Plot

```{r fig.height=5.5, fig.width=6, message=FALSE, warning=FALSE}
plot(qnorm((173+ 1:173+0.5)/(2*173+ 1.125)),sort(abs(res.p2)),
     xlab= 'Expected Half-Normal Order Stats', ylab= 'Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)   
```

Conclusion: 
* This one captured a few points at the end, most of points fall around the reference line.
* As the linear deviation revealed that the overdispersion that leads to lack of fit.

(3) Change the Model: Adjusting for Overdispersion

```{r}
## Adjusting for Overdispersion: phi parameter
dispersion.M = summary(crab.M2, dispersion = phi)
```

The new model is here after adjusted for over-dispersion:
$$log(\hat {Sa}_i) = 0.44744W_t + 0.04590 * W -1.29168$$
Interpretation: 
* The coefficients stay the same as before the adjustment for over-dispersion, while for the Std. Error it changes after adding up the phi parameter.
* Since estimate of $\beta_1 = 0.44744$> 0, for one unit of increase in the female crab weight, the expected number of male satellites will increase, and will be multiplied by $e^{0.44744}$ = 1.5643 times.

* Although estimate of $\beta_2 = 0.045$ is also positive, the coefficient didn't past the Z-test which indicates that there is no significant influence of female crab width on the expected number of male satellites.

#### Goodness of fit: Overdispersion

```{r}
## Methods 1: Genearalized Pearson Chi-square
res.p1 = residuals(crab.M2, type = 'pearson',data = crab)  
G2 = sum(res.p1^2) # calc dispersion param based on full model
## [1] 536.5963

df2 = 173 - 3
1 - pchisq(G2, df = df2) ## 
## df = 170, p-value = 0 
```

The Generalized chi-square test results show us that the p-value is 0 < 0.05, we can reject the null-hypothesis, meaning that there is overdispersion in the model.

### Problem 2

#### a) Fit a poisson model 

```{r message=FALSE, warning=FALSE}
library(MASS)
parasite_raw = read.table('./HW5-parasite.txt', header = T)
parasite_raw = parasite_raw %>% 
  mutate(Area = as.factor(Area),
         Year = as.factor(Year) )

##test missing values in N. 
colSums(is.na(parasite_raw)) 
parasite_raw = na.omit(parasite_raw)

ps.glm = glm(Intensity ~ Length + Area + Year, data = parasite_raw, family = poisson(link = log))
summary(ps.glm)
```

$$log(\hat{Parasites}) = 2.6432 -0.0284 * Length -0.21195 * Area2 -0.1169 * Area3 +  1.4049 * Area4 + 0.6703 * Year2000 -0.2181 * Year2001$$

Interpretation: As all coefficients have significant p-values with them, we can have interpretations as,

* Intercept: For fish being in Year 1999 and Area 1 with zero Length, the expected number of parasites is exp(2.6432) = 14.057.

* For a fish in certain Area and year, with one unit increase of **Length** of fish, the expected number of parasites will significantly decrease, and will be multiplied by $e^{\beta_1}$ = 0.9719 times.

* Being in **Area2** significantly decreases the expected number of parasites for fish when holding other variables constant, and will be multiplied by $e^{\beta_2}$ = 0.8090 times.

* Being in **Area3** significantly decreases the expected number of parasites for fish when holding other variables constant, and will be multiplied by $e^{\beta_3}$ = 0.8897 times.

* Being in **Area4** significantly increases the expected number of parasites for fish when holding other variables constant, and will be multiplied by $e^{\beta_4}$ = 4.075 times.

* Being in **Year2000** significantly increases the expected number of parasites for fish when holding other variables constant, and will be multiplied by $e^{\beta_5}$ = 1.9547 times.

* Being in **Year2001** decreases the expected number of parasites for fish when holding other variables constant, and will be multiplied by $e^{\beta_6}$ =  0.8040 times.


### b). Goodness of fit 

```{r}
## Genearalized Pearson Chi-square
res.p3 = residuals(ps.glm, type = 'pearson')  
G3 = sum(res.p3^2) # calc dispersion param based on full model
G3
df3 = (1254 - 63) - (4 - 1) - (3 - 1) - 1  ## (n - missing) - p 

1 - pchisq(G3, df = df3)

## Alternative Methods: Deviance 
D3 = deviance(ps.glm)

1 - pchisq(D3, df = df3) 
```

Conclusions: 

Based on the generalized pearson chi-square statistics and Deviance methods, we find that we cannot reject the null hypothesis, meaning that this model fits well.


### c). 0-inflation model 

Thus, the zip model has two parts, a poisson count model and the logit model for predicting excess zeros, both contained three predictors for modelling. 
```{r}
## Test zeros in dataset.
nrow(subset(parasite_raw, parasite_raw$Intensity == 0)) 

m1 <- zeroinfl( Intensity ~  Length + Area + Year, data = parasite_raw) 
# child and camper for poisson, persons for binary
summary(m1)
```

#### Model processes

In these 2 processes both give zeros:
* True zero: fish is not susceptible to parasites, captured in the binary model; 
* psedo zero: fish is susceptible to parasites but have no parasite, captured in the poisson model;

Interpretation of *zero-inflation model* coefficients: 

* Being in **Area2** increases the odds of being susceptible to parasites by 105% (exp(0.7186)= 2.0517), and this is statistically significant (p = 0.00015 < 0.05);

* Being in **Area3** increases the odds of being susceptible to parasites by 93% (exp(0.6577) = 1.9303), and this is statistically significant (p = 8.53e-05 < 0.05);

* Being in **Area4** decreases the odds of being susceptible to parasites by 64.04% (exp(-1.0229) = 0.3596, 1 - 0.3596 = 0.6404), and this is statistically significant (p = 5.48e-08 < 0.05);

* Being in **Year2000** decreases the odds of having the opportunity of susceptible to parasites by 52.86% (exp(-0.7521)= 0.4714, 1 - 0.4714 = 0.5286), and this is statistically significant (p = 1.37e-05 < 0.05);

* Being in **Year2001** increases the odds of having the opportunity of susceptible to parasites by 57.86% (exp(0.4565) = 1.5786), and this is statistically significant (p = 0.00152 < 0.05);

* With one unit increase of fish **length**, the odds of being susceptible strains to parasites decreases by 0.98% (exp(-0.009889) = 0.9901, 1-0.9901 = 0.0098).


Interpretation of *poisson model* coefficients: 

* Among the fish species who are susceptible to parasites, The expected rate of parasites intensity the fish had for one unit increase of **length** is 0.9638 (exp(-0.0368067) = 0.9638), holding other variables constant, and this is statistically significant (p<.0001); 

* Among the fish species who are susceptible to parasites, being in **Area2** increases the expected rate of parasites intensity by 30.85% (exp(0.2689) = 1.3085), holding other variables constant, and this is statistically significant
(p<.0001). 

* Among the fish species who are susceptible to parasites, being in **Area3** increases the expected rate of parasites intensity by 93.03% (exp(0.1463) = 1.9303), holding other variables constant, and this is statistically significant
(p<.0001);

* Among the fish species who are susceptible to parasites, being in **Area4** significantly increases the expected rate of parasites by 157% (exp(0.9448) = 2.5722); 

* Among the fish species who are susceptible to parasites, being in **year 2000** increases the expected rate of parasites by 47.98% (exp(0.3919) = 1.4798);

* Among the fish species who are susceptible to parasites, being in **year 2001** decreases the expected rate of parasites intensity by 4.38% (exp(-0.0448) = 0.9562, 1- 0.9562 = 0.0438).




